Compositional equivalences based on Open pNets

### 2020-04-21 2016-Weak-FH-Bisim/WeakBisim.tex
lecture part 1

       	Let $\mathcal{R}''$ be the relation: 
       	\[\mathcal{R}'' = 
       	\{(s,u|\Pred_{s,u})\,\,\Big|\,\Pred_{s,u}=\bigvee_{\begin{array}{c}       		
       		(s,t|\Pred_{s,t})\in\mathcal{R}\\ (t,u|\Pred_{t,u})\in\mathcal{R}' 	
       		\end{array}
       	}\,\Pred_{s,t}\land\Pred_{t,u}\}\]

@@

       	First, we have $\forall y, \forall x, \forall p,  J_1=J_{px}=J_{pxy}, 
       	({s}',u_{pxy}|\Pred\,'_{{pxy}})\in \mathcal{R}''$ for some $\Pred\,'_{{pxy}}$. 
       	Indeed for any 
       	$p$, 
       	$x$, and 
       	$y$, $t_{px}$
       	relates ${s}'$ and $u_{pxy}$, we have
       	$(s',t_{px}|\Pred_{{px}})\in \mathcal{R}$
       	and $(t_{px},u_{pxy}|\Pred_{{pxy}})\in \mathcal{R}'$. 
       	More precisely,  $t_{px} \in ({t'_p})^{p\in P'}$ where $({t'_p})^{p\in 
       		P'}$ is 
       	the set of states relating ${s}'$ and $u_{pxy}$ (the states used in the open transition must belong to the set of states ensuring the transitive relation).
       	Additionally, for all $p$, $x$, $y$, $\Pred_{px}\land 
       	\Pred_{{pxy}}\implies 
       	\Pred\,'_{{pxy}}$ (this is one element of the  disjunction defining the 
       	predicate $\Pred\,'_{{pxy}}$
       	relating ${s}'$ and $u_{pxy}$ in the definition of $\mathcal{R}''$).

One can notice that, as bisimulation predicates are used to relate states that 
belong to two different open automata, the free variables of these predicates 
that do not belong to the two related automata can safely be renamed to avoid any 
name clash. In practice,
we can suppose that, as $\Pred\,'_{{pxy}}$ is used to relate states of 
$\mathcal{T}_1$, and $\mathcal{T}_3$, there exists a predicate obtained by 
renaming variables, that is equivalent to $\Pred\,'_{{pxy}}$ and does not 
contain the variables of 	$\mathcal{T}_2$. This predicate with ``fresh'' 
variables is called   $\Pred\,'_{{pxy}}$ in the following.
Similarly, we can suppose that $\Pred_{{px}}$ contains no 
variable in $\mathcal{T}_3$, and $\Pred_{{pxy}}$ contains no 
variable in $\mathcal{T}_1$.

       	Second, by definition of bisimulation we need (recall that $\Pred_{s,u}$ is the 
       	original predicate relating $s$ and $u$ by definition of the transitive 
       	closure):\\
       	\begin{small}
       		$\Pred_{s,u} \land \Pred_{OT}\\
       		\hspace{1cm} \implies\!\!\! \displaystyle{\bigvee_{x\in X}\bigvee_{y\in Y}\bigvee_{p\in P}
       		\left( \forall j. \beta_j=\beta_{jpxy}  \land \Pred_{OT_{pxy}}
       		\land \alpha\!=\!\alpha_{pxy} \land
       		\Pred\,'_{{pxy}}\subst{\Post_{OT}\uplus\Post_{OT_{pxy}}}\right)}$.
       	\end{small}
From (*) and (**) we have, for all $p$
       	\begin{small}
       		$\Pred_{s,p} \land \Pred_{OT}\land\Pred_{p,u}\\
       		\hspace{1cm} \implies\!\!\! \displaystyle{\bigvee_{x\in X}
       		\left( \forall j. \beta_j=\beta_{jpx}  \land \Pred_{OT_{px}}
       		\land \alpha\!=\!\alpha_{px} \land
       		\Pred_{{px}}\subst{\Post_{OT}\uplus\Post_{OT_{px}}}\right)\land\Pred_{p,u}}\\
       		%
       		\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}
       		\left( \forall j. \beta_j=\beta_{jpx}  \land 
       		(\Pred_{OT_{px}}\land\Pred_{p,u})
       		\land \alpha\!=\!\alpha_{px} \land
       		\Pred_{{px}}\subst{\Post_{OT}\uplus\Post_{OT_{px}}}\right)\\
       		%
       		\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}
       		\big( \forall j. \beta_j=\beta_{jpx}  \land (\bigvee_{y\in Y} 
       		\big( \forall j'. \beta_{j'px}=\beta_{j'pxy}  \land \Pred_{OT_{pxy}}
       		\land \alpha_{px}\!=\!\alpha_{pxy}\\~\hspace{3em}~ \land
       		\Pred_{{pxy}}\subst{\Post_{OT_{px}}\uplus\Post_{OT_{pxy}}}\big))
       		\land \alpha\!=\!\alpha_{px} \land
       		\Pred_{{px}}\subst{\Post_{OT}\uplus\Post_{OT_{px}}}\big)\\
       		%
       		\hspace{1cm} \implies\!\!\! \bigvee_{x\in X} \bigvee_{y\in Y}
       		\big( \forall j, j'. \beta_j=\beta_{jpx} \land \beta_{j'px}=\beta_{j'pxy}
       		\land \big( 
       		\Pred_{OT_{pxy}}
       		\land \alpha\!=\alpha_{px}\!=\!\alpha_{pxy}\\~\hspace{3.5em}~ \land
       		\Pred_{{pxy}}\subst{\Post_{OT_{px}}\uplus\Post_{OT_{pxy}}}
       		\land
       		\Pred_{{px}}\subst{\Post_{OT}\uplus\Post_{OT_{px}}}\big)\big)
       		$

@@

By construction, four substitutions $\subst{~}$ only have an effect on the  
variables of the open automaton they belong to, they also produce terms containing only 
variables of the open automaton they belong to. Finally, because of the domain of the 
substitutions  and of the predicates, we have:\\
       	$\Pred_{{px}}\subst{\Post_{OT}\uplus\Post_{OT_{px}}}\land 
       	\Pred_{{pxy}}\subst{\Post_{OT_{px}}\uplus\Post_{OT_{pxy}}}\\
       	\Leftrightarrow % we changed =
       	\Pred_{{px}}\subst{\Post_{OT}\uplus\Post_{OT_{px}}\uplus\Post_{OT_{pxy}}}\land
       	\Pred_{{pxy}}\subst{\Post_{OT}\uplus\Post_{OT_{px}}\uplus\Post_{OT_{pxy}}}\\
       	\implies\Pred\,'_{{pxy}}\subst{\Post_{OT}\uplus\Post_{OT_{px}}\uplus\Post_{OT_{pxy}}}\\
       	\Leftrightarrow % we changed = 
       	 \Pred\,'_{{pxy}}\subst{\Post_{OT}\uplus\Post_{OT_{pxy}}} $\\

@@

      	\begin{small}     	
$\displaystyle{\Pred_{s,u} \land \Pred_{OT}}\\
 \implies\!\!\! \bigvee_{p\in P} (
	\Pred_{s,p}\land\Pred_{p,u} \land \Pred_{OT})\\
\hspace{1cm} \implies\!\!\!\bigvee_{p\in P}
 \bigvee_{x\in X} \bigvee_{y\in Y}
\big( \forall j, j'.\beta_j=\beta_{jpx} \land \beta_{j'px}=\beta_{j'pxy}
\land\\ ~\hspace{7em} 
\Pred_{OT_{pxy}}
\land \alpha\!=\!\alpha_{px}\!=\!\alpha_{pxy}\land\Pred\,'_{{pxy}}\subst{\Post_{OT}\uplus \Post_{OT_{pxy}}}\big)
\\
\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}\bigvee_{y\in Y}\bigvee_{p\in P}
\left( \forall j. \beta_j\!=\!\beta_{jpxy}  \land \Pred_{OT_{pxy}}
\land \alpha\!=\!\alpha_{pxy} \land
\Pred\,'_{{pxy}}\subst{\Post_{OT}\uplus\Post_{OT_{pxy}}}\right)$
  \end{small}

@@

\subsubsection*{{\bf Lemma~\ref{lem-decompose}}: Open transition decomposition} ~\\
 Consider two pNets $P$ and $Q$ that are not pLTSs\footnote{A similar lemma can be proven for a pLTS $Q$}.
	Let $\Leaves(Q)=p_l^{l\in L_Q}$; suppose:
	\[ P[Q]_{j_0}  
		\models
		{\openrule
			{
				\beta_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $\exists i\in L_Q.\,s_i\neq s'_i$, i.e. $Q$ takes part in the reduction.
		 Then there exist $\alpha_Q$, $\Pred\,'$, $\Pred\,''$, 
		$\Post\,'$, $\Post\,''$ s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\openrule
			{
				\beta_j^{j\in (J\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTarrow {\alpha}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
	\vspace{-2.2ex}\\\text{and~~}
		%\[
		Q\models{\openrule
			{
				\beta_j^{j\in J\cap\Holes(Q)}, \Pred\,'',  
				\Post\,''}
			{\ostate{s_i^{i\in L_Q}} \OTarrow {\alpha_Q}
				\ostate{s_i'^{\,i\in L_Q}}}
		}%\]
		\end{mathpar}
		and  $\Pred \iff \Pred\,'
		\land \Pred\,''\land \alpha_Q=\beta_{j_0}$, $\Post=\Post\,'\uplus 
		\Post\,''$ where $\Post\,''$ is the restriction of $\Post$ over variables of 
		$\Leaves(Q)$.

   \begin{proof}
Consider rule \TrDeux\ in 
       Definition~\ref{def:operationalSemantics}, applied to the pNet $P[Q]_{j_0}$. 	
	\noindent
\begin{small}	
\begin{mathpar}
    \mprset {vskip=.8ex}
\inferrule
    {
\Leaves(\mylangle {\pNet}_m^{m\in I}, \set{\Sort}, \symb{SV}_k^{k\in 
    	K}\myrangle) \!=\! \pLTS_l^{l\in L} \qquad  	
k\!\in\! K \qquad SV_k \!=\! \SV{(\alpha'_m)^{m \in I_1\uplus I_2\uplus J}}{\alpha'}{e_b} 
\\
\\     	
	\forall m\!\!\in\!\! I_1. {\pNet_m 
	\models\openrule
    	{
    	\beta_{j}^{j\in J_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {\alpha_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}} }	
  \qquad
\forall m\!\!\in\!\! I_2.		{ \pNet_m 
    	 \models
    	\openrule
    	{\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {\alpha_m}
    		\ostate{s_m'}} }\\\\
     J' = \biguplus_{m\in I_1}\!\! J_m \uplus J 	\\
    	\Pred = \bigwedge_{m\in I_1\uplus I_2}\!\! \Pred_m \land
    	\Predsv(SV_k,\alpha_m^{m\in I_1\uplus I_2},\beta_j^{j\in J},\alpha)\\ 
    	\forall i\in	L\backslash \left(\biguplus_{m\in I_1}\!\! L_m \uplus I_2\right).\,s'_i=s_i \\
    \fresh(\alpha'_m,\alpha',\beta_j,\alpha) 
    }
    {\mylangle {\pNet}_m^{m\in I}, \set{\Sort}, \symb{SV}_k^{k\in K}\myrangle
    	\models
    	{\openrule
    		{
    		{\beta_j}^{j\in J^\prime}, \Pred,  \biguplus_{m\in I_1\uplus I_2} 
    		\Post_m}
    		{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
    			\ostate{(s_i^\prime)^{i\in L}}}
    	}
    }\quad\TrDeux
\end{mathpar}

@@

Now focus on OTs\footnote{In the appendix we use OT to abbreviate ``open transition'' and WOT to stand for ``weak open transition.''} of the subnets. For each $m\in I_1\uplus I_2$ we have one of the two 
following OT\footnotemark[\thefootnote]:\\
for $m$ in $I_1$
\[
\pNet_m \models\openrule
    	{
    	\beta_{j}^{j\in J_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {\alpha_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}}\]
Or, $m$ in $I_2$
\[{ \pNet_m 
    	 \models
    	\openrule
    	{\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {\alpha_m}
    		\ostate{s_m'}}}\]

Only elements of $(I_1\uplus I_2)\backslash\{j_0\}$ and  are useful to assert the premise for reduction of $P$; the last 
one ensures the open transition for the pNet $Q$ (note that $Q$ is at place $j_0$, and by 
definition of the open transition 
for $P[Q]_{j_0}$, 
$L_{j_0}=L_Q$, and $J_{j_0}=	J\cap\Holes(Q)$):\\[-2ex]
	\[Q\models{\openrule
		{
			\beta_j^{j\in J\cap\Holes(Q)}, \Pred_{j_0},  
			\Post\,''}
		{\ostate{s_i^{i\in L_Q}} \OTarrow {\alpha_{j_0}}
			\ostate{(s_i')^{\, i\in L_Q}}}
	}\]

>>>>

       	Let $\mathcal{R}''$ be the relation: 
       	\[\mathcal{R}'' = 
       	\{(s,u|\Pred_{s,u})\,\,\Big|\,\Pred_{s,u}=\hspace{-0.7cm}\bigvee_{\begin{array}{c}       		
       		(s,t|\Pred_{s,t})\in\mathcal{R}\\ (t,u|\Pred_{t,u})\in\mathcal{R}' 	
       		\end{array}
       }\hspace{-0.5cm}\Pred_{s,t}\land\Pred_{t,u}\}\]

@@

       	First, we have $\forall y, \forall x, \forall p,  J_1=J_{px}=J_{pxy}, 
       	({s}',u_{pxy}|\Pred\,'_{{pxy}})\in \mathcal{R}''$ for some $\Pred\,'_{{pxy}}$. 
       	Indeed for any 
       	$p$, 
       	$x$, and 
       	$y$, $t_{px}$
       	relates ${s}'$ and $u_{pxy}$, we have
       	$(s',t_{px}|\Pred_{{px}})\in \mathcal{R}$
       	and $(t_{px},u_{pxy}|\Pred_{{pxy}})\in \mathcal{R}'$. 
       	More precisely,  $t_{px} \in ({t'_p})^{p\in P'}$ where $({t'_p})^{p\in 
       		P'}$\hl{$P'\subseteq P$} is 
       	the set of states relating ${s}'$ and $u_{pxy}$ (the states used in the open transition must belong to the set of states ensuring the transitive relation).
       	Additionally, for all $p$, $x$, $y$, $\Pred_{px}\land 
       	\Pred_{{pxy}}\implies 
       	\Pred\,'_{{pxy}}$ (this is one element of the  disjunction defining the 
       	predicate $\Pred\,'_{{pxy}}$
       	relating ${s}'$ and $u_{pxy}$ in the definition of $\mathcal{R}''$).

One can notice that, as bisimulation predicates are used to relate states that 
belong to two different open automata, the free variables of these predicates 
that do not belong to the two related automata can safely be renamed to avoid any 
name clash. In practice,
we can suppose that, as $\Pred\,'_{{pxy}}$ is used to relate states of 
$\mathcal{T}_1$ and $\mathcal{T}_3$, there exists a predicate obtained by 
renaming variables, that is equivalent to $\Pred\,'_{{pxy}}$ and does not 
contain the variables of 	$\mathcal{T}_2$. This \hl{predicate with ``fresh'' 
variables is called}   $\Pred\,'_{{pxy}}$ in the following.
Similarly, we can suppose that $\Pred_{{px}}$ contains no 
variable in $\mathcal{T}_3$, and $\Pred_{{pxy}}$ contains no 
variable in $\mathcal{T}_1$.

       	Second, by definition of bisimulation we need (recall that $\Pred_{s,u}$ is the 
       	original predicate relating $s$ and $u$ by definition of the transitive 
       	closure):\\
       	\begin{small}
       		$\Pred_{s,u} \land \Pred_{OT}\\
       		\hspace{1cm} \implies\!\!\! \displaystyle{\bigvee_{x\in X}\bigvee_{y\in Y}\bigvee_{p\in P}
       		\left( \forall j. \beta_j=\beta_{jpxy}  \land \Pred_{OT_{pxy}}
       		\land \alpha\!=\!\alpha_{pxy} \land
       		\Pred\,'_{{pxy}}\subst{\Post_{OT}\uplus\Post_{OT_{pxy}}}\right)}$.
       	\end{small}

From (*) and (**) we have, for all $p$
       	\begin{small}
       		$\Pred_{s,p} \land \Pred_{OT}\land\Pred_{p,u}\\
       		\hspace{1cm} \implies\!\!\! \displaystyle{\bigvee_{x\in X}
       		\left( \forall j. \beta_j=\beta_{jpx}  \land \Pred_{OT_{px}}
       		\land \alpha\!=\!\alpha_{px} \land
       		\Pred_{{px}}\subst{\Post_{OT}\uplus\Post_{OT_{px}}}\right)\land\Pred_{p,u}}\\
       		%
       		\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}
       		\left( \forall j. \beta_j=\beta_{jpx}  \land 
       		(\Pred_{OT_{px}}\land\Pred_{p,u})
       		\land \alpha\!=\!\alpha_{px} \land
       		\Pred_{{px}}\subst{\Post_{OT}\uplus\Post_{OT_{px}}}\right)\\
       		%
       		\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}
       		\big( \forall j. \beta_j=\beta_{jpx}  \land (\bigvee_{y\in Y} 
       		\big( \forall j'. \beta_{j'px}=\beta_{j'pxy}  \land \Pred_{OT_{pxy}}
       		\land \alpha_{px}\!=\!\alpha_{pxy}\\~\hspace*{2em}~ \land
       		\Pred_{{pxy}}\subst{\Post_{OT_{px}}\uplus\Post_{OT_{pxy}}}\big))
       		\land \alpha\!=\!\alpha_{px} \land
       		\Pred_{{px}}\subst{\Post_{OT}\uplus\Post_{OT_{px}}}\big)\\
       		%
       		\hspace{1cm} \implies\!\!\! \bigvee_{x\in X} \bigvee_{y\in Y}
       		\big( \forall j, j'. \beta_j=\beta_{jpx} \land \beta_{j'px}=\beta_{j'pxy}
       		\land \big( 
       		\Pred_{OT_{pxy}}
       		\land \alpha\!=\alpha_{px}\!=\!\alpha_{pxy}\\~\hspace*{2em}~ \land
       		\Pred_{{pxy}}\subst{\Post_{OT_{px}}\uplus\Post_{OT_{pxy}}}
       		\land
       		\Pred_{{px}}\subst{\Post_{OT}\uplus\Post_{OT_{px}}}\big)\big)
       		$

@@

By construction, four substitutions $\subst{~}$ only have an effect on the  
variables of the open automaton they belong to, they also produce terms containing only 
variables of the open automaton they belong to. Finally, because of the domain of the 
substitutions of the predicates, we have:\\
       	$\Pred_{{px}}\subst{\Post_{OT}\uplus\Post_{OT_{px}}}\land 
       	\Pred_{{pxy}}\subst{\Post_{OT_{px}}\uplus
       	\Post_{OT_{pxy}}}\Leftrightarrow\\ % we changed =
       	\Pred_{{px}}\subst{\Post_{OT}\uplus\Post_{OT_{px}}\uplus\Post_{OT_{pxy}}}\land
       	\Pred_{{pxy}}\subst{\Post_{OT}\uplus\Post_{OT_{px}}\uplus\Post_{OT_{pxy}}}\\
       	\implies\Pred\,'_{{pxy}}\subst{\Post_{OT}\uplus
       	\Post_{OT_{px}}\uplus\Post_{OT_{pxy}}}
       	\Leftrightarrow\\ % we changed = 
       	 \Pred\,'_{{pxy}}\subst{\Post_{OT}\uplus\Post_{OT_{pxy}}} $\\

@@

      	\begin{small}     	
$\displaystyle{\Pred_{s,u} \land \Pred_{OT}}\\
 \implies\!\!\! \bigvee_{p\in P} (
	\Pred_{s,p}\land\Pred_{p,u} \land \Pred_{OT})\\
\hspace{1cm} \implies\!\!\!\bigvee_{p\in P}
 \bigvee_{x\in X} \bigvee_{y\in Y}
\big( \forall j, j'.\beta_j=\beta_{jpx} \land \beta_{j'px}=\beta_{j'pxy}
\land \Pred_{OT_{pxy}}\\
~\hspace*{2em}~\land \alpha\!=\!\alpha_{px}\!=\!\alpha_{pxy}\land\Pred\,'_{{pxy}}\subst{\Post_{OT}\uplus \Post_{OT_{pxy}}}\big)
\\
\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}\bigvee_{y\in Y}\bigvee_{p\in P}
\left( \forall j. \beta_j\!=\!\beta_{jpxy}  \land \Pred_{OT_{pxy}}
\land \alpha\!=\!\alpha_{pxy} \land
\Pred\,'_{{pxy}}\subst{\Post_{OT}\uplus\Post_{OT_{pxy}}}\right)$
  \end{small}

@@

\subsubsection*{{\bf Lemma~\ref{lem-decompose}}: Open transition decomposition} ~\\
 Consider two pNets $P$ and $Q$ that are not pLTSs\footnote{A similar lemma can be proven for a pLTS $Q$}.
	Let $\Leaves(Q)=p_l^{l\in L_Q}$; suppose:
	\[ P[Q]_{j_0}  
		\models
		{\openrule
			{
				\beta_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $\exists i\in L_Q.\,s_i\neq s'_i$, i.e. $Q$ takes part in the reduction.
		 Then, there exist $\alpha_Q$, $\Pred\,'$, $\Pred\,''$, 
		$\Post\,'$, $\Post\,''$ s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\openrule
			{
				\beta_j^{j\in (J\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTarrow {\alpha}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
	\vspace{-2.2ex}\\\text{and~~}
		%\[
		Q\models{\openrule
			{
				\beta_j^{j\in J\cap\Holes(Q)}, \Pred\,'',  
				\Post\,''}
			{\ostate{s_i^{i\in L_Q}} \OTarrow {\alpha_Q}
				\ostate{s_i'^{\,i\in L_Q}}}
		}%\]
		\end{mathpar}
		and  $\Pred \iff \Pred\,'
		\land \Pred\,''\land \alpha_Q=\beta_{j_0}$, $\Post=\Post\,'\uplus 
		\Post\,''$ where $\Post\,''$ is the restriction of $\Post$ over variables of 
		$\Leaves(Q)$.

   \begin{proof}
Consider rule \TrDeux\ in 
       Definition~\ref{def:operationalSemantics}, applied to the pNet $P[Q]_{j_0}$. 	
	\noindent
\begin{small}	
\begin{mathpar}
    \mprset {vskip=.8ex}
\inferrule
    {
\Leaves(\mylangle {\pNet}_m^{m\in I}, \set{\Sort}, \symb{SV}_k^{k\in 
    	K}\myrangle) \!=\! \pLTS_l^{l\in L} \qquad  	
k\!\in\! K \qquad SV_k \!=\! \SV{(\alpha'_m)^{m \in I_1\uplus I_2\uplus J}}{\alpha'}{e_b} 
\\
\\     	
	\forall m\!\in\! I_1. {\pNet_m 
	\models\openrule
    	{
    	\beta_{j}^{j\in J_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {\alpha_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}} }	
  \qquad
\forall m\!\in\! I_2.		{ \pNet_m 
    	 \models
    	\openrule
    	{\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {\alpha_m}
    		\ostate{s_m'}} }\\\\
     J' = \biguplus_{m\in I_1}\!\! J_m \uplus J 	\\
    	\Pred = \bigwedge_{m\in I_1\uplus I_2}\!\! \Pred_m \land
    	\Predsv(SV_k,\alpha_m^{m\in I_1\uplus I_2},\beta_j^{j\in J},\alpha)\\ 
    	\forall i\in	L\backslash \left(\biguplus_{m\in I_1}\!\! L_m \uplus I_2\right).\,s'_i=s_i \\
    \fresh(\alpha'_m,\alpha',\beta_j,\alpha) 
    }
    {\mylangle {\pNet}_m^{m\in I}, \set{\Sort}, \symb{SV}_k^{k\in K}\myrangle
    	\models
    	{\openrule
    		{
    		{\beta_j}^{j\in J^\prime}, \Pred,  \biguplus_{m\in I_1\uplus I_2} 
    		\Post_m}
    		{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
    			\ostate{(s_i^\prime)^{i\in L}}}
    	}
    }\quad\TrDeux
\end{mathpar}

@@

\hl{Now focus on OTs}\footnote{In the appendix we use OT to abbreviate ``open transition'' and WOT to stand for ``weak open transition.''} of the subnets. For each $m\in I_1\uplus I_2$ we have one of the two 
following OT\footnotemark[\thefootnote]:\\
either $m$ in $I_1$
\[
\pNet_m \models\openrule
    	{
    	\beta_{j}^{j\in J_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {\alpha_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}}\]
or, $m$ in $I_2$
\[{ \pNet_m 
    	 \models
    	\openrule
    	{\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {\alpha_m}
    		\ostate{s_m'}}}\]

\hl{Only elements of} $(I_1\uplus I_2)\backslash\{j_0\}$ and  are useful to assert the premise for reduction of $P$; the last 
one ensures the open transition for the pNet $Q$ (note that $Q$ is at place $j_0$, and by 
definition of the open transition 
for $P[Q]_{j_0}$, 
$L_{j_0}=L_Q$, and $J_{j_0}=	J\cap\Holes(Q)$):\\[-2ex]
	\[Q\models{\openrule
		{
			\beta_j^{j\in J\cap\Holes(Q)}, \Pred_{j_0},  
			\Post\,''}
		{\ostate{s_i^{i\in L_Q}} \OTarrow {\alpha_{j_0}}
			\ostate{(s_i')^{\, i\in L_Q}}}
	}\]
\TODO{j'ai du mal, je reviens}

This already ensures the second part of the conclusion of the lemma, i.e. the OT for $Q$ 
if we 
choose\footnotemark[\thefootnote]  $\alpha_Q=\alpha_{j_0}$ and $\Pred\,''= \Pred_{j_0}$. 
Considering 
the OT of $P$ we have another  $J'$ that is $J'_p=J'\setminus\Holes(Q)\uplus 
\{j_0\}$; we denote $I'_1=I_1\setminus \{j_0\}$ the predicate is 
$\displaystyle{\Pred\,'=\!\!\!\!\!\bigwedge_{m\in I_1'\uplus I_2}\!\!\!\!\!\Pred_m  \land \Predsv(SV_k,\alpha_m^{m \in I'_1\uplus I_2},\beta_j^{j\in J\cup \{j_0\}},\alpha)}$
where\footnotemark[\thefootnote] $\Predsv(SV_k,\alpha_m^{m \in I'_1\uplus I_2},\beta_j^{j\in 
J\cup 
\{j_0\}},\alpha)\Leftrightarrow 
\forall i\in I'_1\uplus I_2.\, \alpha_i=\alpha'_i\land \forall j \in J\cup \{j_0\}.\, 
\beta_j=\alpha'_j 
\land 
\alpha=\alpha'
\land e_b$. Modulo renaming of fresh variables, this is identical to the predicate that 
occurs in 
the source open transition except $\alpha_{j_0}=\alpha'_{j_0}$ has been replaced by  
$\beta_{j_0}=\alpha'_{j_0}$. As $\alpha_{j_0}=\alpha_Q$ and $\beta_{j_0}$ is free, we 
have $\beta_{j_0}=\alpha'_{j_0}\land \beta_{j_0}=\alpha_Q \iff 
\alpha_{j_0}=\alpha'_{j_0}$.
Thus, $\Pred \iff (\Pred\,'
		\land \Pred\,'')\land \alpha_Q=\beta_{j_0}$. 
Finally, Post 
%conditions are easily split\footnote{\TODO{moved. could
%    even remove...?}Note that if the rules are not simplified, no
%  post-condition range over the variables of $P$ and $Q$ at the same
%  time} 
into conditions of the context $P$ and the pNet $Q$ (they are
builts similarly as they only deal with  
leaves): $\Post=\Post\,'\uplus \Post\,''$. This concludes the 
proof as we checked all the premises of the open transition for both $P$ and $Q$. We obtain the following reduction by the rule \TrDeux
	\noindent
	\begin{small}
\begin{mathpar}
    \mprset {vskip=.8ex}
\inferrule
    {
\Leaves(\mylangle {\pNet}_m^{m\in I\setminus\{j_0\}}, \set{\Sort}, \symb{SV}_k^{k\in 
    	K}\myrangle) \!=\! \pLTS_l^{l\in L} \qquad  	
k\!\in\! K \qquad SV_k \!=\! \SV{(\alpha'_m)^{m \in I_1\uplus I_2\uplus J}}{\alpha'}{e_b} 
\\
\\     	
	\forall m\!\!\in\!\! I_1\setminus\{j_0\}. {\pNet_m 
	\models\openrule
    	{
    	\beta_{j}^{j\in J_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {\alpha_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}} }	
  \qquad
\forall m\!\!\in\!\! I_2.		{ \pNet_m 
    	 \models
    	\openrule
    	{\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {\alpha_m}
    		\ostate{s_m'}} }\\\\
     J' = \biguplus_{m\in I_1 \setminus \{j_0\}}\!\! J_m \uplus J 	\\
    	\Pred\,' = \bigwedge_{m\in I_1\uplus I_2\setminus\{j_0\}}\!\! \Pred_m \land
    	\Predsv(SV_k,\alpha_m^{m\in I_1\uplus I_2 \setminus \{j_0\}},\beta_j^{j\in J\cup\{j_0\}},\alpha)\\ 
    	\forall i\in	L\backslash \left(\biguplus_{m\in I_1\setminus\{j_0\}}\!\! L_m \uplus I_2\right).\,s'_i=s_i \\
    \fresh(\alpha'_m,\alpha',\beta_j,\alpha) 
    }
    {\mylangle {\pNet}_m^{m\in I\setminus\{j_0\}}, \set{\Sort}, \symb{SV}_k^{k\in K}\myrangle
    	\models
    	{\openrule
    		{
    		{\beta_j}^{j\in J\setminus \Holes(Q)\uplus \{j_0\}}, \Pred\,',  \biguplus_{m\in I_1\setminus\{j_0\}\uplus I_2} 
    		\Post_m}
    		{\ostate{s_i^{i\in L\setminus L_Q}} \OTarrow {\alpha}
    			\ostate{(s_i^\prime)^{i\in L\setminus L_Q}}}
    	}
    }
\end{mathpar}
 \end{small}
\qed

\footnotetext{Cloning and freshness introduce alpha-conversion at many points 
of the proof; we 
	only 
	give major arguments concerning alpha-conversion to make the proof readable; in 
	general, fresh variables appear in each transition inside
        terms $\beta_j$, $v$, and 
	$\Pred$.}
\end{proof}
